---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clust431

<!-- badges: start -->
<!-- badges: end -->

The goal of clust431 is to preform hierarchical clustering and k-means clustering.

## Installation

You can install the released version of clust431 from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("clust431")
```

```{r echo = T, include = F}
library(clust431)
library(tidyverse)
library(sna)
```

## Example K-Means

K-Means clustering example on a dataset of two variables using k_means

```{r}
mydata <- select(iris, Petal.Length, Petal.Width)


k_means(mydata, 3, PCA = F)
```

Comparison with kmeans function in the stats package

```{r}
kmeans(mydata, 3)
```

Example of k_means using PCA 

```{r}
k_means(mtcars, 3, PCA = T)
```

Example of k_means using PCA and the k-means++ algorithm for improved initialization
```{r}
k_means(mtcars, 3, PCA = T, plus.plus = T)
```


## Example Hierarchical Clustering

Hierarchical clustering on a dataset

```{r}
mydata <- select(mtcars, mpg, cyl) %>% head(6)

hier_clust(mydata)
```

Comparison using the hclust function in the stats package
```{r}
mydata %>% dist() %>% hclust() %>% plot(sub = "")
```

Example of hierachical clustering using the manhattan distance metric
```{r}
hier_clust(mydata, method = 'manhattan')
```

Example of hierachical clustering using a large dataset 
```{r}
hier_clust(mtcars)
```
